---
title: "Complement in IBD"
format: 
  html:
    code-fold: true
    toc: true
  pdf:
    toc: true
    number-sections: true
    colorlinks: true
    keep-tex: true
    include-in-header: 
       text: |
         \usepackage{fvextra}
         \DefineVerbatimEnvironment{Highlighting}{Verbatim}{breaklines,commandchars=\\\{\}}
include-before-body:
  text: |
    \RecustomVerbatimEnvironment{verbatim}{Verbatim}{
      showspaces = false,
      showtabs = false,
      breaksymbolleft={},
      breaklines
      % Note: setting commandchars=\\\{\} here will cause an error 
    }
editor_options: 
  chunk_output_type: console
execute:
  cache: true

---

This project analysed the interaction between complement system components and intestinal microbiota in the context of intestinal inflammation in IBD. Data acquisition and experiments were conducted from 02-2016 until 2018 in Innsbruck, Austria.

```{r setup}
#| warning: false
library(tidyverse)
library(rstatix)
library(ggpubr)
library(reshape2)
library(dunn.test)
library(RColorBrewer)
library(colorspace)
source("code/FunctionsCustomBoxplotStat.R")
source("code/FunctionsHeatmapPlotting.R")

options(scipen = 999) 

my_cols <- c("#669999", "#cc0000")
cutoffval = round(log(50), 2)       # sets hard limit on range of heatmap at 50 fold expression (for log heatmaps)


# read datasets
dat.qpcr = read.csv("data/qpcr_expression-data.csv", stringsAsFactors = T) %>%
  mutate(DISEASE = factor(DISEASE, labels = c("CD", "Control", "UC")),
         SITE = factor(SITE, labels = c("Ileum", "Cecum", "Ascending colon", "Transverse Colon", "Descending Colon", "Sigmoid", "Rectum")))
dat.elisa = read.csv("data/cohortB_ELISA.csv", sep = ";")
dat.fc = read.csv("data/cohortB_FC.csv", sep = ";")
dat.cohortB.demographics = read.csv("data/cohortB_Demographics.csv", sep =";")
dat.cohortB.lab = read.csv("data/CohortB_Lab.csv", sep = ";")
dat.wieslab = read.csv("data/cohortB_Wieslab.csv", sep = ";")

# remove empty rows 
dat.elisa = filter(dat.elisa, !is.na(PROT_C3))%>%
  left_join(dat.cohortB.demographics, by = "ID") %>%
  left_join(dat.cohortB.lab, by = "ID")
dat.wieslab = filter(dat.wieslab, !is.na(WIESLAB_CL1)) %>%
  left_join(dat.cohortB.demographics, by = "ID") %>%
  left_join(dat.cohortB.lab, by = "ID")
dat.fc = dat.fc %>%
  left_join(dat.cohortB.demographics, by = "ID") %>%
  left_join(dat.cohortB.lab, by = "ID") %>%
  filter(!is.na(PERC_IGA), DIS != "uncl") %>%
  mutate(SEX = factor(SEX, levels = c("male", "female"), labels = c("male", "female")),
         DIS = factor(DIS, levels = c("cd", "contr", "uc"), labels = c("CD", "Healthy", "UC")),
         DIS = relevel(DIS, ref = "Healthy"),
         CALPRO2 = cut(CALPRO, breaks = c(-Inf, 100, Inf)),
         CALPRO4 = cut(CALPRO, breaks = c(-Inf, 50, 250, Inf),labels = c("< 50", "50-250", "> 250")),
         M_A = ifelse(M_A %in% c(1,2,3), M_A, NA) %>% factor(levels = c(1,2,3), labels = c("A1", "A2", "A3")),
         M_B = ifelse(M_B %in% c(1,2,3,4), M_B, NA) %>% factor(levels = c(1,2,3,4), labels = c("B1","B2","B3", "B3p")),
         M_L = ifelse(M_L %in% c(1,2,3,4), M_L, NA) %>% factor(levels = c(1,2,3,4), labels = c("L1", "L2", "L3", "L4")),
         M_E = ifelse(M_E %in% c(1,2,3), M_E, NA) %>% factor(levels = c(1,2,3), labels = c("E1", "E2", "E3"))) 

# sorting data are already an .rds file containing a phylosec object generated by a separate script
# dat.sort = readRDS("data/dada2physeq.RDS")
```

# Background information and Introduction

The complement project includes 3 distinct cohorts:

1.  Complement_CohortA_2023-02-17.xlsx
2.  Complement_CohortB_2023-02-17.xlsx
3.  Complement_CohortSorting_2023-02-17.xlsx

Cohort A contains patients for which intestinal biopsies were collected and therefore is the basis of qPCR analysis.

Cohort B is a subcohort of the Innsbruck IBD cohort biobank which includes fecal samples. From this cohort, Flow Cytometry measurements and protein levels in feces and serum are derived.

Cohort C is another small subcohort of the Innsbruck IBD cohort. Fecal samples from these patients have been subjected to fluorescence assisted cell sorting (FACS). From these data, microbiome composition has been derived.

For data analysis, separate *.csv* files for each data types were exported from the relevant excel files and are available from the *data/* subfolder.

## Introduction

# Methods

# Results

## Cohort statistics

include information on composition of all three subcohorts regarding the following levels:

-   disease diagnosis

-   sex

-   lesional-non-lesional

For qPCR: include Histo and Site

for CohortB: ...

for Cohort C: MEan calpro levels etc.

### Cohort A: qPCR

```{r cohort_statA}
# Cohort statistics of the qPCR subcohort

summary(dat.qpcr$DISEASE) %>% t() %>%
  as.data.frame() %>%
  pivot_longer(CD:UC,names_to = "Disease", values_to = "n") %>%
  knitr::kable(caption = "patient numbers by disease in qpcr cohort")

  
table(dat.qpcr$LESIONAL, dat.qpcr$DISEASE) %>%
  knitr::kable(caption = "Number of lesional and non-lesional samples per disease subcohort")

table(dat.qpcr$SITE, dat.qpcr$DISEASE) %>% 
    knitr::kable(caption = "Number of samples per biopsy site and disease subcohort")

```

### Cohort B: ELISA

```{r cohort_statB_ELISA}
table(dat.elisa$SEX, dat.elisa$DIS)
```

### Cohort B: Wieslab

### Cohort B: Flow Cytometry

### Microbiome Sorting

## Expression of complement proteins in the intestine of patients with IBD and healthy controls

Data from qPCR analysis will be presented as a heatmap and also as table with median + IQR. First, factor variables are reformatted for better presentation in the final heatmap. Next, the number of missing values per target is calculated and stored to file.

```{r qpcr_descriptive}

# Reformat factor columns
dat.qpcr$DISEASE <- factor(dat.qpcr$DISEASE, levels = c("Control","CD","UC"), labels = c("Control", "CD", "UC"))
dat.qpcr$LESIONAL <- factor(dat.qpcr$LESIONAL, levels = c("NL", "L"), labels = c("Non Lesional", "Lesional") )
dat.qpcr$HISTO <- factor(dat.qpcr$HISTO, levels = c(0:3), labels = c("No Inflammation", "Mild Inflammation", "Moderate Inflammation", "Severe Inflammation"))
targets <- names(dat.qpcr[7:ncol(dat.qpcr)]) # save names of qPCR targets into 1 variable
dat.qpcr$grouping <- paste(dat.qpcr$DISEASE, dat.qpcr$LESIONAL) # grouping variables for significance testings
dat.qpcr$grouping<-relevel(factor(dat.qpcr$grouping), "Control Non Lesional")  #relevel so that healthy is first

# Number of missing values for PCR results
na_count <-sapply(dat.qpcr, function(y) sum(length(which(is.na(y))))) %>% 
  data.frame() %>% 
  rownames_to_column("variable") %>%
  filter(variable %in% targets)
names(na_count) = c("Variable", "NA_Count") 
write_excel_csv(na_count, file = "results/qpcr/qpcr_targets_nacount.csv")

knitr::kable(na_count, caption = "Number of missing values per qPCR target")

#### Descriptive Statistics ####
desctable <- dat.qpcr %>%
  mutate(Group = paste(DISEASE, LESIONAL)) %>%
  select(-ID, -SITE, -HISTO, -grouping, - DISEASE, - LESIONAL)%>%
  gather(key = Target, value = value, C1q_alpha:VSIG4) %>%
  mutate(Group = relevel(as_factor(Group), ref ="Control Non Lesional")) %>%
  group_by(Target, Group) %>%
  summarize(
    count = n(),
    mean = mean(value, na.rm = TRUE),
    sd = sd(value, na.rm = TRUE),
    median = median(value, na.rm = TRUE),
    min = min(value, na.rm = TRUE),
    max = max(value, na.rm = TRUE),
    IQR = IQR(value, na.rm =TRUE)
  )

desctable2<- dat.qpcr %>%
  mutate(DISEASE=ifelse(DISEASE != "Control", "IBD", "Control"),
         Group = paste(DISEASE, LESIONAL)) %>%
  select(-ID, -SITE, -HISTO, -grouping, - DISEASE, - LESIONAL) %>%
  gather(key = Target, value = value, C1q_alpha:VSIG4) %>%
  mutate(Group = relevel(as_factor(Group), ref ="Control Non Lesional")) %>%
  group_by(Target, Group) %>%
  summarize(
    count = n(),
    mean = mean(value, na.rm = TRUE),
    sd = sd(value, na.rm = TRUE),
    median = median(value, na.rm = TRUE),
    min = min(value, na.rm = TRUE),
    max = max(value, na.rm = TRUE),
    IQR = IQR(value, na.rm =TRUE)
  )

```

For the publication, median and IQR values will be used because these data are not normally distributed. The table is rearranged to present these values and then saved to file:

```{r qpcr_descriptive_pubtable}
qpcr.descriptive1 = desctable %>%
  mutate(val = paste(round(median, 2), "±", round(IQR, 2))) %>%
  select(Target, Group, val) %>%
  pivot_wider(names_from = Group, values_from = val)

qpcr.descriptive2 = desctable2 %>%
  mutate(val = paste(round(median, 2), "±", round(IQR, 2))) %>%
  select(Target, Group, val) %>%
  pivot_wider(names_from = Group, values_from = val)

write_excel_csv(desctable, file = "results/qpcr/descriptivestats_uccd.csv")
write_excel_csv(desctable2, file = "results/qpcr/descriptivestats.csv")
write_excel_csv(qpcr.descriptive1, file = "results/qpcr/descriptivestats_uccd_publication.csv")
write_excel_csv(qpcr.descriptive2, file = "results/qpcr/descriptivestats_publication.csv")

knitr::kable(qpcr.descriptive1, caption = "Descriptive statistics of relative expression by disease entity and lesional status [median ± IQR]")

knitr::kable(qpcr.descriptive2, caption = "Descriptive statistics of relative expression by disease status (healthy vs IBD) and lesional status [median ± IQR]")
```

### Boxplots of relative expression

Next, relative expression is shown using boxplots for all targets. Both Control vs IBD and Control vs CD vs UC comparisons are performed:

```{r qpcr_boxplots}
#| warning: false
#| fig-height: 11
#| fig-width: 9
  
# drop unused columns
datbp <- select(dat.qpcr, -c(HISTO, SITE, grouping))
#melt datbp
datbpm <- melt(datbp, id.vars = c("ID", "LESIONAL", "DISEASE"))
levels(datbpm$DISEASE) = c("Control", "CD", "UC")
datbpm$DISEASE2 <-ifelse(datbpm$DISEASE == "Control", "Control", "IBD") %>% factor()


boxplot_healthy_ibd <- ggplot(datbpm, aes(x = DISEASE2, y = value)) + 
  stat_summary(aes(fill = LESIONAL), alpha = 0.8,
               position = position_dodge(preserve ="single"),
               fun.data = calc_boxplot_stat, geom = "boxplot") +
  labs(title = "Median ddCT", fill = "",x = "", y = "Expression (qPCR)")+
  scale_y_continuous() +
  theme_classic() +
    scale_fill_manual(values = my_cols) +
  facet_wrap(~variable, scales  = "free") +
  expand_limits(x = 0, y = 0) +
  theme(panel.spacing.y = unit(0.5, "cm"))

boxplot_disease <- ggplot(datbpm, aes(x = DISEASE, y = value)) + 
  stat_summary(aes(fill = LESIONAL), alpha = 0.8,
               position = position_dodge(preserve ="single"),
               fun.data = calc_boxplot_stat, geom = "boxplot")+
  labs(title = "Median ddCT", fill = "",x = "", y = "Expression (qPCR)")+
  scale_y_continuous() +
  theme_classic() +
  scale_fill_manual(values = my_cols) +
  facet_wrap(~variable, scales  = "free") +
  theme(panel.spacing.y = unit(0.5, "cm"))

ggsave(boxplot_healthy_ibd, filename = "results/qpcr/boxplot_ControlIBD.png", width = 8.2, height = 9)
ggsave(boxplot_healthy_ibd, filename = "results/qpcr/boxplot_ControlIBD.pdf", width = 8.2, height = 9)
ggsave(boxplot_disease, filename = "results/qpcr/boxplot_ControlCDUC.png", width = 8.5, height = 9)
ggsave(boxplot_disease, filename = "results/qpcr/boxplot_ControlCDUC.pdf", width = 8.7, height = 9)

boxplot_healthy_ibd
boxplot_disease
```

### Heatmaps

Heatmaps are plotted using separate helper functions. These are defined in a R script file which is stored in the *code/* folder.

```{r qpcr_heatmap}
#| fig-height: 5
#| fig-width: 5
  
dat.m <- melt(dat.qpcr, id.vars=c("ID", "LESIONAL", "DISEASE", "SITE"))

## subsetting the original df, dropping site and histo, then calculating column mean and median by Disease and Lesion status --> 5 groups!
dat.means <- select(dat.qpcr, DISEASE, LESIONAL, all_of(targets)) %>% group_by(DISEASE, LESIONAL) %>% summarise_all(tibble::lst(mean), na.rm = TRUE)       
dat.medians <- select(dat.qpcr, DISEASE, LESIONAL, all_of(targets)) %>% group_by(DISEASE, LESIONAL) %>% summarise_all(tibble::lst(median), na.rm = TRUE)       

## Format Dataframes for heatmap        
heatmap_means <- prepare_df_hmp(datafr = dat.means, cutoffval = cutoffval)
heatmap_medians <- prepare_df_hmp(datafr = dat.medians, cutoffval = cutoffval)


## Plot both heatmaps and save to file
plot_heatmap(datafr = heatmap_medians, caption_string = "Macroscopic Aspect - Medians", file_suffix = "medians", dpival = 300)
plot_heatmap(datafr = heatmap_means, caption_string = "Macroscopic Aspect - Means", file_suffix = "means", dpival = 300)
```

Finally, statistical significance is assessed for all targets comparing lesional vs non-lesional sites within diseases and non-lesional sites from patients with controls.

```{r qpcr_statistics}
kruskalloop <- function(targets_list, n) {
  f <- as.formula(paste(targets_list[n], "~", "grouping"))
  kruskal.test(f, data = dat.qpcr)
}

kruskallist = list()

for (i in seq_along(targets)) {
  y = kruskalloop(targets,i)
  kruskallist[[targets[i]]] <- y        
}

kruskaltbl<-do.call(rbind, kruskallist)
kruskaltbl_print <- as.data.frame(kruskaltbl) %>%
  select(statistic, parameter, p.value, data.name) %>%
  rownames_to_column("Target") %>%
  select(Target, "p value" = p.value, "degrees of freedom" = parameter, "chi squared" = statistic)

wilcoxlist_bonf = list()
wilcoxlist_bonf2 = list()

dat.qpcr$grouping2 = ifelse(dat.qpcr$DISEASE=="Control", "Control", "IBD")
dat.qpcr$grouping2 = paste0(dat.qpcr$grouping2, ".", dat.qpcr$LESIONAL)

wilcoxloop_bonf <- function(targets_list, n) {
  f <- eval(parse(text = (paste0("dat.qpcr$", targets_list[n]))))
  pairwise.wilcox.test(f,dat.qpcr$grouping, p.adjust = "bonf", na.rm=T)
}

wilcoxloop_bonf2 <- function(targets_list, n) {
  f <- eval(parse(text = (paste0("dat.qpcr$", targets_list[n]))))
  pairwise.wilcox.test(f,dat.qpcr$grouping2, p.adjust = "bonf", na.rm=T)
}


for (i in seq_along(targets)) {
  wilcoxy = wilcoxloop_bonf(targets, i)
  wilcoxlist_bonf[[i]] <- wilcoxy
  wilcoxlist_bonf[[i]]$Target <- targets[i]
}

for (i in seq_along(targets)) {
  wilcoxy = wilcoxloop_bonf2(targets, i)
  wilcoxlist_bonf2[[i]] <- wilcoxy
  wilcoxlist_bonf2[[i]]$Target <- targets[i]
}

results_bonf <- do.call(rbind, wilcoxlist_bonf)
results_bonf2 <- do.call(rbind, wilcoxlist_bonf2)

bonf_resultlist = list()
bonf2_resultlist = list()

for (i in seq_along(targets)) {
  x <- unnest(as.data.frame(results_bonf), "Target")
  tab <- x %>%
    filter(Target == targets[i])
  tab<-as.data.frame(tab$p.value)
  bonf_resultlist[[targets[i]]] = tab
}   

for (i in seq_along(targets)) {
  x <- unnest(as.data.frame(results_bonf2), "Target")
  tab <- x %>%
    filter(Target == targets[i])
  tab<-as.data.frame(tab$p.value)
  bonf2_resultlist[[targets[i]]] = tab
}   

bonf_finalres = do.call(rbind, bonf_resultlist) %>% as.data.frame() %>%
  rownames_to_column("Variable") %>%
  mutate(Variable = str_replace(Variable, "[.]", "__")) %>%
  separate(Variable, into = c("Target", "Comparison"), sep = "__")
bonf2_finalres = do.call(rbind, bonf2_resultlist) %>% as.data.frame() %>%
  rownames_to_column("Variable") %>%
  mutate(Variable = str_replace(Variable, "[.]", "__")) %>%
  separate(Variable, into = c("Target", "Comparison"), sep = "__")

significance_ControlIBD = bonf2_finalres %>%
  filter(if_any(c(Control.Non.Lesional, IBD.Lesional), ~. < 0.05))
         
significance_ControlCDUC = bonf_finalres %>%
  filter(if_any(Control.Non.Lesional:UC.Lesional, ~. < 0.05))

# Write final Bonferroni corrected p-values to file
write_excel_csv(significance_ControlIBD, file = "results/qpcr/ControlVsIBD_pairwise-wilcox_bonferroni_filtered.csv")
write_excel_csv(significance_ControlCDUC, file = "results/qpcr/ControlVsCDUC_pairwise-wilcox_bonferroni_filtered.csv")

write_excel_csv(bonf2_finalres, file = "results/qpcr/ControlVsIBD_pairwise-wilcox_bonferroni.csv")
write_excel_csv(bonf_finalres, file = "results/qpcr/ControlVsCDUC_pairwise-wilcox_bonferroni.csv")

knitr::kable(significance_ControlIBD, digits = 2,
             caption = "Significantly differently expressed qPCR targets by disease status (pairwise wilcoxon tests")

knitr::kable(significance_ControlCDUC, digits = 2,
             caption = "Significantly differently expressed qPCR targets by disease entity (pairwise wilcoxon tests")

```

## Complement protein levels in feces and blood of patients with IBD

elisa and wieslab

## Complement opsonisation of intestinal microbiota

Bacteria were isolated from fecal samples using centrifugation and washing steps and stained for complement and antibody opsonisation using the following antibodies:

***Insert antibody table here!***

| Default | Left | Right | Center |
|---------|:-----|------:|:------:|
| 12      | 12   |    12 |   12   |
| 123     | 123  |   123 |  123   |
| 1       | 1    |     1 |   1    |

: Antibodies used for flow cytometry {#tbl-flowcyt_antibodies}

see @tbl-flowcyt_antibodies

### Opsonisation is increased in IBD
First, a summary statistics table for all flow cytometry derived parameters is generated (this includes the percentage of events single positive for IgA or C3c as well as the share of double positive and double negative events). The results are shown in @tbl-flow_cyt_summarystats.

```{r tbl-flow_cyt_summarystats}
#| warning: false
#| tbl-cap: "Summary Statistics of flow cytometry derived parameters of opsonisation"
#### Per Disease ####
fcdat_summarystat = dat.fc %>%
  select(DIS, PERC_IGA:MEAN_IGA) %>%
  group_by(DIS) %>%
  summarise(across(everything(), list(mean = mean, sd = sd, median= median, IQR = IQR), 
                   na.rm =T, .names = "{col}__{fn}")) %>%
  tidyr::pivot_longer(cols = c(PERC_IGA__mean:MEAN_IGA__IQR), names_to = "var", values_to = "value") %>%
  tidyr::separate(var, into = c("Variable", "Stat"), sep ="__") %>%
  tidyr::pivot_wider(id_cols = c(Variable, DIS), names_from = Stat, values_from = value) %>%
  arrange(Variable, DIS)

knitr::kable(fcdat_summarystat, digits = 2, col.names = c("Variable", "Disease Group", "Mean", "Std.Dev", "Median", "IQR"))
```

@tbl-flow_cyt_wilcoxon shows a significant increase in IgA opsonisation (in terms of % opsonized bacteria). Complement opsonisation is increased in CD patients only. In these, also the median amount of complement deposited on bacterial surfaces is higher than in healthy controls. This effect is not detected for IgA opsonisation.

```{r tbl-flow_cyt_wilcoxon}
#| tbl-cap: "Results of pairwise wilcoxon tests on flow cytometry derived parameters"
# Significance testing per variable
stat.perc_iga = dat.fc %>%
  dplyr::select(DIS, PERC_IGA) %>%
  pairwise_wilcox_test(PERC_IGA ~ DIS, p.adjust.method = "bonferroni") %>%
  mutate(p_round = round(p.adj, digits = 3),
         p_round = ifelse(p_round < 0.001, "< 0.001", p_round)) 
stat.perc_c3c = dat.fc %>%
  dplyr::select(DIS, PERC_C3C) %>%
  pairwise_wilcox_test(PERC_C3C ~ DIS, p.adjust.method = "bonferroni") %>%
  mutate(p_round = round(p.adj, digits = 3),
         p_round = ifelse(p_round < 0.001, "< 0.001", p_round))
stat.perc_dblep = dat.fc %>%
  dplyr::select(DIS, PERC_DBLEP) %>%
  pairwise_wilcox_test(PERC_DBLEP ~ DIS, p.adjust.method = "bonferroni") %>%
  mutate(p_round = round(p.adj, digits = 3),
         p_round = ifelse(p_round < 0.001, "< 0.001", p_round))
stat.perc_dblen = dat.fc %>%
  dplyr::select(DIS, PERC_DBLEN) %>%
  pairwise_wilcox_test(PERC_DBLEN ~ DIS, p.adjust.method = "bonferroni") %>%
  mutate(p_round = round(p.adj, digits = 3),
         p_round = ifelse(p_round < 0.001, "< 0.001", p_round))
stat.med_iga = dat.fc %>%
  dplyr::select(DIS, MED_IGA) %>%
  pairwise_wilcox_test(MED_IGA ~ DIS, p.adjust.method = "bonferroni") %>%
  mutate(p_round = round(p.adj, digits = 3),
         p_round = ifelse(p_round < 0.001, "< 0.001", p_round))

stat.med_c3c = dat.fc %>%
  dplyr::select(DIS, MED_C3C) %>%
  pairwise_wilcox_test(MED_C3C ~ DIS, p.adjust.method = "bonferroni") %>%
  mutate(p_round = round(p.adj, digits = 3),
         p_round = ifelse(p_round < 0.001, "< 0.001", p_round))

wilcoxon_all = rbind(stat.perc_iga, stat.perc_c3c, stat.perc_dblep, stat.perc_dblen,
                     stat.med_iga, stat.med_c3c) %>%
  select(Variable = ".y.", group1, group2, n1,n2,statistic, p, p.adj = "p_round", p.adj.signif)

knitr::kable(wilcoxon_all, digits = 2, col.names = c("Variable", "Group1", "Group2", "n1", "n2", "Test statistic", "p value", "Bonferroni adj. p value", "Significance"))
```

Next, we generate figures to illustrate this finding (@fig-flow_cyt_percentages shows the percentages of single and double positive events per disease and @fig-flow_cyt_medians depicts median fluorescence intensities for IgA and C3c per disease).

```{r fig-flow_cyt_percentages}
#| fig-cap: "Percentages of IgA, C3c, double pos. and double neg. flow cytometry events per disease"
#| fig-height: 6
#| fig-width: 6

p.perc_iga = ggboxplot(dat.fc, x = "DIS", y = "PERC_IGA",
                       colour = "DIS", fill = "DIS", 
                       add = "jitter", alpha = 0.5) +
  stat_pvalue_manual(stat.perc_iga, label ="p.adj.signif", y.position = c(90, 100, 110), tip.length = 0.01)+ 
  theme_classic(base_size = 11) +
  theme(legend.position = "none") +
  labs(x = "", y = "% of IgA pos. events") +
  scale_fill_brewer(type = "qual", palette = 2)

p.perc_c3c = ggboxplot(filter(dat.fc, PERC_C3C < 20), x = "DIS", y = "PERC_C3C",
                       colour = "DIS", fill = "DIS", 
                       add = "jitter", alpha = 0.5) +
  stat_pvalue_manual(stat.perc_c3c, label ="p.adj.signif", y.position = c(2,2.5,3), tip.length = 0.01) +
  theme_classic(base_size = 11) +
  theme(legend.position = "none") +
  labs(x = "", y = "% of C3c pos. events") +
  scale_fill_brewer(type = "qual", palette = 2)

p.perc_dblep = ggboxplot(dat.fc, x = "DIS", y = "PERC_DBLEP",
                       colour = "DIS", fill = "DIS", 
                       add = "jitter", alpha = 0.5) +
  stat_pvalue_manual(stat.perc_dblep, label ="p.adj.signif", y.position = c(18,20,22), tip.length = 0.01) +
  theme_classic(base_size = 11) +
  theme(legend.position = "none") +
  labs(x = "", y = "% of double pos. events") +
  scale_fill_brewer(type = "qual", palette = 2)

p.perc_dblen = ggboxplot(dat.fc, x = "DIS", y = "PERC_DBLEN",
                       colour = "DIS", fill = "DIS", 
                       add = "jitter", alpha = 0.5) +
  stat_pvalue_manual(stat.perc_dblen, label ="p.adj.signif", y.position = c(100,106,112), tip.length = 0.01) +
  theme_classic(base_size = 11) +
  theme(legend.position = "none") +
  labs(x = "", y = "% of double neg. events") +
  scale_fill_brewer(type = "qual", palette = 2)


percentages4_disease = cowplot::plot_grid(p.perc_iga, p.perc_c3c, p.perc_dblep, p.perc_dblen,align = "hv", labels = c("A","B", "C", "D"))
percentages2_disease = cowplot::plot_grid(p.perc_iga, p.perc_c3c,align = "hv", labels = c("A","B"))

# Save plots for publication
ggsave(plot = percentages4_disease, filename = "results/flow/PerDisease_Percentages.pdf", device = "pdf", width = 6, height = 6)
ggsave(plot = percentages2_disease, filename = "results/flow/PerDisease_Percentages_noDoubles.pdf", device = "pdf", width = 6, height = 3)

percentages4_disease
```

```{r fig-flow_cyt_medians}
#| fig-cap: "Median fluorescence intensity for IgA and C3c per disease"
#| fig-height: 3
#| fig-width: 6
options(scipen = 0)
p.med_iga = ggboxplot(dat.fc, x = "DIS", y = "MED_IGA",
                       colour = "DIS", fill = "DIS", 
                       add = "jitter", alpha = 0.5) +
  stat_pvalue_manual(stat.med_iga, label ="p.adj.signif", y.position = c(8.2e09, 10e09, 9.2e09), tip.length = 0.01)+ 
  theme_classic(base_size = 11) +
  theme(legend.position = "none") +
  labs(x = "", y = "Median IgA-APC fluorescence") +
  scale_fill_brewer(type = "qual", palette = 2)

p.med_c3c = ggboxplot(dat.fc, x = "DIS", y = "MED_C3C",
                       colour = "DIS", fill = "DIS", 
                       add = "jitter", alpha = 0.5) +
  stat_pvalue_manual(stat.med_c3c, label ="p.adj.signif", y.position = c(6e09, 7e09, 8e09), tip.length = 0.01)+ 
  theme_classic(base_size = 11) +
  theme(legend.position = "none") +
  labs(x = "", y = "Median C3c-FITC fluorescence") +
  scale_fill_brewer(type = "qual", palette = 2)
  
perdisease_median = cowplot::plot_grid(p.med_iga, p.med_c3c, labels = c("A", "B"), align = "hv")
# save plot for publication
ggsave(plot = perdisease_median, filename = "results/flow/PerDisease_MedianF.pdf", device = "pdf", width = 6, height = 3)

perdisease_median

```


### Microbiota opsonisation by IgA is correlated to fecal calprotectin
IgA opsonisation of microbiota  correlates with intestinal inflammation. Fecal calprotectin is split into ranges from < 50 µg/g, 50-250 µg/g and > 250 µg/g, reflecting clinically relevant cutoffs for healthy, quiescent and active IBD. Summary statistics are outlined in @tbl-flow_cyt-summarystat-percalpro. Statistical significance testing (wilcoxon tests) is reported in @tbl-flow_cyt-wilcoxon_percalpro. Finally, @tbl-flow_cyt-correlation-calpro shows results of a correlation analysis with a continual calprotectin variable.

```{r tbl-flow_cyt-summarystat-percalpro}
#| warning: false
#| tbl-cap: "Summary Statistics of flow cytometry derived parameters of opsonisation related to calprotectin range"
options(scipen=999)
fcdat_summarystat_calpro4 = dat.fc %>%
  select(DIS, PERC_IGA:MEAN_IGA, CALPRO4) %>%
  filter(!is.na(CALPRO4)) %>%
  group_by(DIS, CALPRO4) %>%
  summarise(across(everything(), list(mean = mean, sd = sd, median= median, IQR = IQR), 
                   na.rm =T, .names = "{col}__{fn}")) %>%
  tidyr::pivot_longer(cols = c(PERC_IGA__mean:MEAN_IGA__IQR), names_to = "var", values_to = "value") %>%
  tidyr::separate(var, into = c("Variable", "Stat"), sep ="__") %>%
  tidyr::pivot_wider(id_cols = c(Variable, DIS, CALPRO4), names_from = Stat, values_from = value) %>%
  arrange(rev(Variable), DIS, CALPRO4)

knitr::kable(fcdat_summarystat_calpro4, digits = 2, col.names = c("Variable", "Disease Group", "Fecal Calprotectin [µg/g]", "Mean", "Std.Dev", "Median", "IQR"))
```

```{r tbl-flow_cyt-wilcoxon_percalpro}
#| tbl-cap: "Results of pairwise wilcoxon tests by fecal calprotectin range"
stat.perc_iga_calpro4 = dat.fc %>%
  dplyr::select(CALPRO4, PERC_IGA) %>%
  pairwise_wilcox_test(PERC_IGA ~ CALPRO4, p.adjust.method = "bonferroni") %>%
  mutate(p_round = round(p.adj, digits = 3),
         p_round = ifelse(p_round < 0.001, "< 0.001", p_round)) %>%
  select(Variable = ".y.", group1, group2, n1,n2,statistic, p, p.adj = "p_round", p.adj.signif)

stat.perc_c3c_calpro4 = dat.fc %>%
  dplyr::select(CALPRO4, PERC_C3C) %>%
  pairwise_wilcox_test(PERC_C3C ~ CALPRO4, p.adjust.method = "bonferroni") %>%
  mutate(p_round = round(p.adj, digits = 3),
         p_round = ifelse(p_round < 0.001, "< 0.001", p_round)) %>%
    select(Variable = ".y.", group1, group2, n1,n2,statistic, p, p.adj = "p_round", p.adj.signif)

stat.med_iga_calpro4 = dat.fc %>%
  dplyr::select(CALPRO4, MED_IGA) %>%
  pairwise_wilcox_test(MED_IGA ~ CALPRO4, p.adjust.method = "bonferroni") %>%
  mutate(p_round = round(p, digits = 3),
         p_round = ifelse(p_round < 0.001, "< 0.001", p_round)) %>%
      select(Variable = ".y.", group1, group2, n1,n2,statistic, p, p.adj = "p_round", p.adj.signif)

stat.med_c3c_calpro4 = dat.fc %>%
  dplyr::select(CALPRO4, MED_C3C) %>%
  pairwise_wilcox_test(MED_C3C ~ CALPRO4, p.adjust.method = "bonferroni") %>%
  mutate(p_round = round(p, digits = 3),
         p_round = ifelse(p_round < 0.001, "< 0.001", p_round)) %>%
      select(Variable = ".y.", group1, group2, n1,n2,statistic, p, p.adj = "p_round", p.adj.signif)

wilcoxon_percalpro_all = rbind(stat.perc_iga_calpro4, stat.perc_c3c_calpro4, stat.med_iga_calpro4, stat.perc_c3c_calpro4)
knitr::kable(wilcoxon_percalpro_all, digits = 2, col.names = c("Variable", "Group1", "Group2", "n1", "n2", "Test statistic", "p value", "Bonferroni adj. p value", "Significance"))

```

```{r tbl-flow_cyt-correlation-calpro}
#| tbl-cap: "Correlation of flow cytometry opsonisation parameters with fecal calprotectin. Results from Spearman correlation analysis"
rbind(cor_test(dat.fc, PERC_IGA, CALPRO, method = "spearman"),
      cor_test(dat.fc, PERC_C3C, CALPRO, method = "spearman"),
      cor_test(dat.fc, MED_IGA, CALPRO, method = "spearman"), 
      cor_test(dat.fc, MED_C3C, CALPRO, method = "spearman")) %>% 
  select(-var2, -statistic, -method) %>%
  mutate(p.signif = ifelse(p< 0.001, "***", ifelse(p<0.01, "**", ifelse(p<0.05, "*", "ns")))) %>%
  knitr::kable(digits = 2, col.names = c("Variable", "Spearman Rho", "p value", "Significance"))

```

@fig-flow_cyt-percentages_calpro shows boxplot on percentages of IgA and C3c positive events and @fig-flow_cyt_medians_calpro depicts boxplots on median fluorescence intensity.

```{r fig-flow_cyt-percentages_calpro}
#| fig-cap: "Percentages of IgA and C3c pos. flow cytometry events per calprotectin range"
#| fig-height: 3
#| fig-width: 6

p.perc_iga_percalpro = ggboxplot(filter(dat.fc, !is.na(CALPRO4)), x = "CALPRO4", y = "PERC_IGA",
                       fill = "CALPRO4", alpha = 0.5, add = "jitter", alpha = 0.5) +
  stat_pvalue_manual(stat.perc_iga_calpro4, label ="p.adj.signif", y.position = c(90, 100, 110), tip.length = 0.01)+ 
  theme_classic(base_size = 11) +
  theme(legend.position = "none") +
  labs(x = "Fecal Calprotectin [µg/g]", y="% of IgA pos. events") +
  scale_fill_brewer(type = "qual", palette = 2)

p.perc_c3c_percalpro = ggboxplot(filter(dat.fc, !is.na(CALPRO4)), x = "CALPRO4", y = "PERC_C3C",
                       fill = "CALPRO4", add = "jitter", alpha = 0.5) +
  stat_pvalue_manual(stat.perc_c3c_calpro4, label ="p.adj.signif", y.position = c(2,2.5,3), tip.length = 0.01)+ 
  theme_classic(base_size = 11) +
  theme(legend.position = "none") +
  labs(x = "Fecal Calprotectin [µg/g]", y="% of C3c pos. events") +
  scale_fill_brewer(type = "qual", palette = 2)

percalpro4_percentage = cowplot::plot_grid(p.perc_iga_percalpro, p.perc_c3c_percalpro, 
                                           labels = c("A", "B"))
percalpro4_percentage

# save to file 
ggsave(filename = "results/flow/PerCalpro4_Percentages_A-B.pdf", device = "pdf",
       plot = percalpro4_percentage, width = 6, height = 3)
```

```{r fig-flow_cyt_medians_calpro}
#| fig-cap: "Median Fluorescence intensities of IgA-APC C3c-FITC per calprotectin range"
#| fig-height: 3
#| fig-width: 6

p.med_iga_percalpro = ggboxplot(filter(dat.fc, !is.na(CALPRO4)), x = "CALPRO4", y = "MED_IGA",
                       fill = "CALPRO4", alpha = 0.5, add = "jitter") +
  stat_pvalue_manual(stat.med_iga_calpro4, label ="p.adj.signif", y.position = c(8.5e09, 9.25e09, 1e10), tip.length = 0.01)+
  theme_classic(base_size = 11) +
  theme(legend.position = "none") +
  labs(x = "Fecal Calprotectin [µg/g]", y="Median IgA-APC fluorescence") +
  scale_fill_brewer(type = "qual", palette = 2)

p.med_c3c_percalpro = ggboxplot(filter(dat.fc, !is.na(CALPRO4)), x = "CALPRO4", y = "MED_C3C",
                       fill = "CALPRO4", alpha = 0.5, add = "jitter") +
  stat_pvalue_manual(stat.med_c3c_calpro4, label ="p.adj.signif", y.position = c(5e09, 6e09, 7e09), tip.length = 0.01)+
  theme_classic(base_size = 11) +
  theme(legend.position = "none") +
  labs(x = "Fecal Calprotectin [µg/g]", y="Median C3c-FITC fluorescence") +
  scale_fill_brewer(type = "qual", palette = 2)

percalpro4_medians = cowplot::plot_grid(p.med_iga_percalpro, p.med_c3c_percalpro, 
                                           labels = c("C", "D"))

percalpro4_medians

# save to file 
ggsave(filename = "results/flow/PerCalpro4_Medians_C-D.pdf", device = "pdf",
       plot = percalpro4_medians, width = 6, height = 3)
```

## Microbiome specificity of complement

sorting data

### Taxonomic data

### Functional mapping

PICRUST

## Microscopy

# Discussion

# Appendix

## References

## Supplementary information
